# Reproducible Research Assignment #1
=======================================

#1 Loading and preprocessing the data
#Load the data
#Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$datetime <- as.POSIXlt(data$date)
data$datetime$min <- data$interval %% 100
data$datetime$hour <- data$interval %/% 100
```

#2 What is mean total number of steps taken per day?

#A: 9354.23
```{r}
stepsPerDay <- tapply(data$steps, data$date, sum, na.rm = T)
averagePerDay <- mean(stepsPerDay)
```
# Make a histogram of the total number of steps taken each day

```{r}
hist(stepsPerDay, xlab = "Number of Steps", main = "Histogram of total number of steps taken each day", col = "red")
```

# Calculate and report the mean and median total number of steps taken per day

```{r}
mean <- mean(stepsPerDay)
median <- median(stepsPerDay)
```

#3 What is the average daily activity pattern?

```{r}
avgStepsPer5min <- tapply(data$steps, data$interval, mean, na.rm = T)
avgStepsPer5min
```

# Make a time series plot of the 5 minute interval (x-axis) and the average number of steps taken, averaged across 
# all days (y-axis) Which 5 minute interval, on average across all the days in the dataset, 
# contains the maximum number of steps?

# A: 835th min

```{r}
plot(names(avgStepsPer5min), avgStepsPer5min, type = 'l', xlab = "5 Minute interval",
     main = "Average number of steps taken, \n averaged across all days",
     ylab = "Number of Steps")
names(which.max(avgStepsPer5min))
```

#4 IMPUTING MISSING VALUES
# 1) Calculate and report the total number of missing values in the dataset

# A: Total number of missing values (NAs) are 2304.

```{r}
totalNA <- sum(is.na(data$steps))
```

# 2) Devise a strategy for filling in all of the missing values in the dataset. The strategy 
#    does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean 
#    for that 5 minute interval, etc.
#   My Strategy: I replaced all the NAs with the median value for that day.

# 3) Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
medianStepsPer5min = tapply(data$steps, data$internal, median, na.rm = T)
correctedData <- data
correctedData[is.na(data$steps), 'steps'] = medianStepsPer5min[as.character(data[is.na(data$steps), 'internal'])]
```

# 4) Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total
#    number of steps taken per day. Do these values differ from the estimates from the first part of the assignment?
#    What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
stepsPerDay = tapply(correctedData$steps, correcedData$date, sum, na.rm = T)
hist(stepsPerday, xlab = "Number of Steps", main = "Histogram of total number of steps taken each day")

mean(stepsPerDay)
median(stepsPerDay)
```

#A: New mean is a bit higher than the old one, and new median is equal to the old one.


# 5 ARE THERE DIFFERENCES IN ACTIVITY PATTERNS BETWEEN WEEKDAYS AND WEEKENDS?
# 1) Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given
# date is a weekday or weekend day.

```{r}
factorValues = as.factor(c('weekday', 'weekend'))
correctedData$dayType = factorValues[correctedData$datetime$wday %in% c(0,6) + 1]
```

# 2) Make a panel plot containing a time series plot of the 5 minute interval (x-axis) and the average number
# of steps taken, averaged across all weekend days or weekend days (y-axis). The plot should look something like the
# following, which was created using simulated data:

```{r}
library(ggplot2)
library(plyr)
avgStepsPer5min = ddply(correctedData, (interval, dayType), summarize, steps = mean(steps))
qplot(interval, steps, data = avgStepsPer5min, facets = dayType ~ ., geom = "line") + ylab("Number of steps") + xlab("Interval")
```
